### 2021-06-25 JSS/using_drtmle_jss.Rmd
minor fix
updates to manuscript. move some sections to appendix

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    address: |
      | 2121 Berkeley Way,
      | Berkeley, California, 94704, U.S.A.
    email: \email{nhejazi@berkeley.edu}
    orcdi: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect. The package additionally implements a targeted inverse probability of
  treatment weighted estimator that yields valid inference even when the
  probability of treatment is estimated via adaptive methods.
header-includes:
  - \usepackage{mathtools}
output: rticles::jss_article
bibliography: refs
---

@@

This has sparked an interest in using adaptive estimation techniques from the
machine learning literature to control for high-dimensional confounders when
estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, standard causal inference
techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa, @robins:2000:epi], fail to yield valid
inference under adaptive estimates of nuisance parameters. More complex
proposals have been developed that *are capable* of utilizing adaptive methods
to control for confounding while still yielding valid inference. These
techniques include augmented inverse probability of treatment weighted
(AIPTW) estimation [@robins:1994:jasa] and targeted minimum loss-based (TML)
estimation [@vdlrubin:2006:ijb]. Under standard assumptions, these estimators
have limiting normal distributions with estimable variance. Estimates of the
asymptotic variance serve as the basis for constructing confidence intervals and
hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of
two key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and
TML estimators a natural appeal: inconsistency in the estimation of one
nuisance parameter may be mitigated by the consistent estimation of the other.
As such, these estimators have increased in popularity in recent years.
However, recent work has shown that the double robustness property does not
necessarily extend to the inferential properties of these estimators when
adaptive estimators of the PS and/or OR are used
[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. \citet{vdl:2014:ijb}
proposed new TML estimators that have an asymptotic normal sampling distribution
with estimable variance under consistent estimation of *either* the PS *or* the
OR, paving the way for inference that is doubly-robust.
\citet{benkeser:2017:biometrika} proposed modifications of these estimators with
similar robustness properties and illustrated their practical performance.

@@

Suppose the observed data consist of $n$ independent and identically
distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
a vector of baseline covariates, $A \in \{0,1\}$ is a binary (for the sake of
simplicity) treatment, $Y$ is an outcome, and $P_0$ is the true data-generating
distribution. The parameters of interest are
\begin{equation}
  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
  \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
covariates is used so that this estimator can be equivalently expressed
\begin{equation*}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
\end{equation*}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator.

@@

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators lack a basis for statistical inference.

@@

\citet{vdl:2014:ijb} demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. \citet{vdl:2014:ijb} derived
modified TML estimators that are doubly-robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly-robust variance estimators that may be used to
construct doubly-robust confidence intervals and hypothesis tests.
\citet{benkeser:2017:biometrika} proposed simplifications of the
\citet{vdl:2014:ijb} estimators and demonstrated the practical performance of
these estimators via simulation experiments.

@@

```{r, cache = TRUE, message = FALSE}
npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
                    stratify = TRUE)
npreg_fit
```

@@

### Arbitrary user-specified regressions

```{r, cache = TRUE, message = FALSE}
# fit a GLM for outcome regression outside of drtmle
out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())

@@

```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
                         targeted_se = FALSE) # needed for this to work
# compare variance to previously obtained
list(glm_fit_uni$drtmle$cov, glm_fit_uni_cv$drtmle$cov)
```

@@

```{r, demo_pcv, cache = TRUE, warning = FALSE}
set.seed(1234)
sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm","SL.mean"),
                     SL_Q = c("SL.glm","SL.mean"),
                     SL_gr = c("SL.glm", "SL.earth"),
                     SL_Qr = c("SL.glm", "SL.earth"),
                     stratify = FALSE,
                     se_cv = "partial")
# compare estimates
# pt estimates should be same
# variance should be larger for pcv
list(
  sl_fit, # no cv
  sl_fit_pcv # partial cv
)
```

@@

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
This confidence interval may be obtained as follows.

@@

```{r, cache = TRUE}
mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
                         glm_gr = "Qn", family = binomial())
mixed_fit_wNAs
```

@@

The user-specified contrasts are also available. For example, we can modify the
`riskRatio` object we used previously for a two-level treatment to compute the
risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.

@@

When considering adaptive estimation techniques, one runs the risk of
overfitting the initial nuisance regressions. While the layer of
cross-validation built into super learning attempts to guard against
overfitting, there are no guarantees in finite samples. To definitively avoid
such overfitting, one can use cross-validated estimates of the nuisance
regressions. Theoretically, the use of cross-validated regression weakens the
assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
This is true of the regular TML and AIPTW estimators, as well as of the TML
estimator with doubly-robust inference. The `drtmle()` function implements
cross-validated estimation of the regression parameters through the `cvFolds`
option, as shown in the code below, where we continue on with the multi-level
treatment example.

```{r, cache = TRUE}
cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_gr = c("SL.glm", "SL.mean"),
                    SL_Qr = c("SL.glm", "SL.mean"),
                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
cv_sl_fit
```

@@

# Additional methods included in drtmle

## Inference for IPTW using adaptive PS estimators

### Implementation of inference for adaptive IPTW

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
options with `drtmle()`; however, as the IPTW estimator does not rely on an
estimate of the OR, options related to the OR are omitted. We demonstrate below,
while continuing with the multi-level treatment example.

```{r, cache = TRUE}
npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
                                   family = binomial(), a_0 = c(0, 1, 2))
npreg_iptw_multiA
```

## Outcome-adaptive propensity score estimation

### Implementation of outcome-adaptive propensity score

```{r, message = FALSE}
# use simplified data
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))

# outcome adaptive propensity score fit using glms
adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
                     adapt_g = TRUE)

@@

>>>>

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    address: |
      | 2121 Berkeley Way,
      | Berkeley, California, 94704, U.S.A.
    email: \email{nhejazi@berkeley.edu}
    orcdi: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect, in addition to several other recently-proposed estimators of the 
  average treatment effect.
header-includes:
  - \usepackage{mathtools}
output: rticles::jss_article
bibliography: refs
---

@@

This has led to increased interest in adopting adaptive estimation techniques from the
machine learning literature to control for high-dimensional confounders when
estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal inference
techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when utilizing
adaptive estimators. In particular, with rare exception, if cross-validation is used
to select tuning parameters for the nuisance parameters, estimates of the population-level
effect will generally be too biased to achieve standard $n^{1/2}$-rate asymptotics. Instead,
*undersmoothing* must be used to select tuning parameters to control the bias 
[@undersmoothed_hal]. However, this can prove quite challenging in practice and 
few guidelines are available.

A more natural framework for integrating adaptive methods into nuisance regression 
estimation are doubly robust approaches, including augmented inverse probability of 
treatment weighted (AIPTW) estimation [@robins:1994:jasa] and targeted minimum 
loss-based (TML) estimation [@vdlrubin:2006:ijb]. The asymptotics of these estimators
are characterized in part by the convergence rates of nuisance parameter estimates with
respect to a proper norm. This fact makes the use of cross-validation more appealing for
selection of tuning parameters for estimators of nuisance parameters, thereby providing 
a more straightforward means of obtaining an estimate of the effect of interest that
enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield closed form
standard error estimates, which can serve as the basis for constructing confidence 
intervals and hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of
two key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and
TML estimators a natural appeal: inconsistency in the estimation of one
nuisance parameter may be mitigated by the consistent estimation of the other.
However, recent work has shown that the double robustness property does not
necessarily extend to the inferential properties of these estimators when
adaptive estimators of the PS and/or OR are used
[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. \citet{vdl:2014:ijb}
proposed new TML estimators that have an asymptotic normal sampling distribution
with estimable variance under consistent estimation of *either* the PS *or* the
OR, paving the way for inference that is doubly-robust.
\citet{benkeser:2017:biometrika} proposed modifications of these estimators with
similar robustness properties and illustrated their practical performance.

@@

Suppose the observed data consist of $n$ independent and identically
distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
a vector of baseline covariates, $A \in \{0,1\}$ is a binary treatment, 
$Y$ is an outcome, and $P_0$ is the true data-generating
distribution. The parameters of interest are
\begin{equation}
  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
  \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
covariates is used so that this estimator can be equivalently expressed
\begin{equation*}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
\end{equation*}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a basis for statistical inference.

@@

\citet{vdl:2014:ijb} demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. \citet{vdl:2014:ijb} derived
modified TML estimators that are doubly-robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly-robust variance estimators that may be used to
construct doubly-robust confidence intervals and hypothesis tests.
\citet{benkeser:2017:biometrika} proposed simplifications of the
\citet{vdl:2014:ijb} estimators and demonstrated the practical performance of
these estimators via simulation.

@@

```{r, cache = TRUE, message = FALSE}
npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
                    stratify = TRUE)
```

In the Appendix, we describe an additional means of using `drtmle()` with 
OR and PS estimates that have been obtained externally.

@@

@@

```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
                         targeted_se = FALSE) # needed for this to work
# compare variance to previously obtained
list(
  `No CV` = glm_fit_uni$drtmle$cov, 
  `CV` = glm_fit_uni_cv$drtmle$cov
)
```

@@

```{r, demo_pcv, cache = TRUE, warning = FALSE}
set.seed(1234)
sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm","SL.mean"),
                     SL_Q = c("SL.glm","SL.mean"),
                     SL_gr = c("SL.glm", "SL.earth"),
                     SL_Qr = c("SL.glm", "SL.earth"),
                     stratify = FALSE,
                     se_cv = "partial")
# compare estimates
# pt estimates should be same
# variance should be larger for pcv
list(
  `No-CV` = sl_fit$cov, # no cv
  `Partial CV` = sl_fit_pcv$cov # partial cv
)
```

@@

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_n(1)}{\psi_n(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
This confidence interval may be obtained as follows.

@@

```{r, cache = TRUE}
mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
                         glm_gr = "Qn", family = binomial())
```

@@

User-specified contrasts are also available. For example, we can modify the
`riskRatio` object we used previously for a two-level treatment to compute the
risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.

@@

To avoid overfitting initial regressions, one can use cross-validated 
estimates of the nuisance
regressions. Theoretically, the use of cross-validated regression weakens the
assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
This is true of the regular TML and AIPTW estimators, as well as of the TML
estimator with doubly-robust inference. The `drtmle()` function implements
cross-validated estimation of the regression parameters through the `cvFolds`
option, as shown in the code below.

```{r, cache = TRUE}
cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_gr = c("SL.glm", "SL.mean"),
                    SL_Qr = c("SL.glm", "SL.mean"),
                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
```

@@

@@

# Appendix

## Arbitrary user-specified regressions

```{r, cache = TRUE, message = FALSE}
# use simplified data
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))

# fit a GLM for outcome regression outside of drtmle
out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())

## Additional methods included in drtmle

### Inference for IPTW using adaptive PS estimators

#### Implementation of inference for adaptive IPTW

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
options with `drtmle()`; however, as the IPTW estimator does not rely on an
estimate of the OR, options related to the OR are omitted.

```{r, cache = TRUE}
npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
                                   family = binomial(), a_0 = c(0, 1))
npreg_iptw_multiA
```

### Outcome-adaptive propensity score estimation

#### Implementation of outcome-adaptive propensity score

```{r, message = FALSE}
# outcome adaptive propensity score fit using glms
adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
                     adapt_g = TRUE)
